Verification of concurrent design patterns with data

### 2019-04-18 2018-Arch-pNets/Coordination Camera ready/arch-pNets.tex
intro + data +open transitions

BIP (Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BliSif08-acp-tc}.
BIP ensures that any property that holds on a BIP
model will also hold on the generated code.
The notion of BIP \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified
%% \footnote{%
%% %
%%   The current approach relies on taxonomies of predefined
%%   architectures.  Future work may include automatic synthesis of
%%   architectures from, \eg CTL~\cite{BaierKatoen2008} formulae.
%% %
%% }
and applied to the model, adding
coordinator components and modifying the 
synchronisation patterns between components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie if two architectures guarantee two properties, their composition ensures the conjunction of the properties.
%
%several architectures are applied, each enforcing a safety
%property, the resulting system satisfies their conjunction.

@@

%\paragraph{Contributions}
The main contributions of this paper are: 1) The definition of a
theory of BIP architectures with data, including a theorem about
preservation of (data dependent) properties by 
compositions. 2) An encoding of architectures with data into  open pNets, allowing for analysis of their temporal properties using pNet's
software tools.  
The  paper is illustrated by a running
example based on the failure monitor architecture from the CubETH
nanosatellite on-board software~\cite{CubETH-case-study}.
%
This running example also relies on the \emph{maximal progress}
assumption, whereby larger interactions are preferred to smaller
ones.  Due to space limitations, we only discuss this
informally. However, proofs of the results provided in the
appendix formally account for maximal progress.

@@

\newCoord{The red dotted rule expresses an implication, stating that if actions of the holes perform the designed actions and the condition is verified, then the variables are modified and the state changes. This implication however occurs in a restricted logic with predefined syntax, at a different logical level compared to  a classical
inference rule expressed in the logic of a scientific paper. 
 The open transition is  a logical implication that uses a simple logic with the 
 expressive power  given by the predicate algebra (it must include logical 
 propositions and equality). 
 Proposition and inference rules of the paper use a general expressive, while predicates inside the open transitions should use a more restricted logic, typically a logic that could  be handled mechanically and expressed by terms that can be encoded in a simple syntax.}

>>>>

BIP (Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BliSif08-acp-tc}.
BIP ensures that any property that holds on a BIP
model will also hold on the generated code.
The notion of BIP \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified
%% \footnote{%
%% %
%%   The current approach relies on taxonomies of predefined
%%   architectures.  Future work may include automatic synthesis of
%%   architectures from, \eg CTL~\cite{BaierKatoen2008} formulae.
%% %
%% }
and applied to the model, adding
coordinator components and modifying the 
synchronisation patterns between components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie if two architectures guarantee two properties, their composition ensures the conjunction of the properties \newCoord{but~\cite{AttieBBJS16-architectures-faoc} did not consider properties depending on data}.
%
%several architectures are applied, each enforcing a safety
%property, the resulting system satisfies their conjunction.

@@

%\paragraph{Contributions}
The main contributions of this paper are: 1) \newCoord{The addition of data to the
theory of BIP architectures, including a theorem about
preservation of data dependent properties by 
compositions.} 2) An encoding of architectures with data into  open pNets, allowing for analysis of their temporal properties using pNet's
software tools.  
The  paper is illustrated by a running
example based on the failure monitor architecture from the CubETH
nanosatellite on-board software~\cite{CubETH-case-study}.
%
This running example also relies on the \emph{maximal progress}
assumption, whereby larger interactions are preferred to smaller
ones.  Due to space limitations, we only discuss this
informally. However, proofs of the results provided in the
appendix formally account for maximal progress.

@@

\newCoord{The red dotted rule somehow expresses an implication, stating that if actions of the holes perform the designed actions and the condition is verified, then the variables are modified and the state changes. This implication however  uses a simple logic with the 
 expressive power  given by the predicate algebra (it must include logical 
 propositions and equality). 
 Proposition and inference rules of the paper use a standard logic, while predicates inside the open transitions should use a more restricted logic, typically a logic that could  be handled mechanically and expressed by terms that can be encoded in a simple syntax.
Open transitions express in a symbolic way, transitions  that are not only parametrised with variables but also actions of not yet known processes.}
